### 2014-05-09 doc/ApTranscriptome_manuscript.Rmd
Incorporated edits to methods from AN

*Aphaenogaster picea* was collected from Molly Bog, Vermont (University of Vermont Natural Areas; lat, lon) and *A. carolinensis* was collected from Durham, North Carolina ( 36.037° N -78.874° W), designated as *A. picea* and *A. carolinensis* henceforth. Though the phylogeny of *Aphaeogaster* is currently unresolved [@umphrey1996], preliminary work supports that these species are separate species (Bernice Bacon DeMarco, pers. communication). species of these species were maintained in the lab for 5 months prior to experimentation.

From a single species of each species, we haphazardly collected 12 ants at the same time on 12 days. Each day, the 12 ants were placed in glass tubes into a water bath (...) at one of 12 randomly-assigned temperatures, every 3.5°C between 0° and 38.5°C, for one hour. The minimum and maximum temperatures were selected based on preliminary work showing that these temperatures are close to the critical minimum and maximum temperatures for an *Aphaenogaster* [@warren2013. At the end of the hour, the ants were flash frozen in liquid nitrogen and stored at -80°C. mRNA was extracted from three pooled ants using a RNeasy kit (Qiagen Inc; Valencia, CA). Ants were homogenized in Buffer ATL with zirconium silicate beads in a Bullet Blender (Next Advance; Averill Park, NY) and the standard Qiagen protocol was followed from this point.

>>>>

Colonies of the ant species *Aphaenogaster picea* were collected from Molly Bog, Vermont (University of Vermont Natural Areas; 44.508°N, -72.702°W) and *Aphaenogaster carolinensis* from Durham, North Carolina ( 36.037° N -78.874° W), designated as *ApVT* and *AcNC* respectively. The phylogeny of *Aphaeogaster* is currently unresolved [@umphrey1996], but putative species identification was possible based on morphology (Bernice Bacon DeMarco, pers. communication). species of these species were maintained in the lab for 6 months prior to experimentation.

From each species, we haphazardly collected 12 ants at the same time on 12 days. Each day, the 12 ants were placed in glass tubes into a water bath at one of 12 randomly-assigned temperatures, every 3.5°C between 0° and 38.5°C, for one hour. The minimum and maximum temperatures were selected based on preliminary work showing that these temperatures are close to the critical minimum and maximum temperatures for *Aphaenogaster* [@warren2013]. At the end of the hour, the ants were flash frozen in liquid nitrogen and stored at -80°C. We extracted mRNA by homogenizing three pooled ants in 500 uL of RNAzol buffer with with zirconium silicate beads in a Bullet Blender (Next Advance; Averill Park, NY), followed by RNAzol extraction (Molecular Research Center Inc; Cincinnati, OH) and then a RNeasy micro extraction (Qiagen Inc; Valencia, CA) following the manufacturer's instructions.
